---
output:
  bookdown::pdf_document2:
    includes:
      in_header: latex/preamble.tex
      before_body: latex/titlepage.tex
    pandoc_args:
    - --csl
    - references/apa.csl
  bookdown::html_document2:
    pandoc_args:
    - --csl
    - references/apa.csl
  bookdown::word_document2:
    pandoc_args:
    - --csl
    - references/apa.csl
toc-title: Table of Contents
bibliography: references/references.bib
link-citations: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0pt}
- \fancyfoot[C]{}
- \fancyfoot[R]{\thepage}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE)
packages <- c("dplyr","readxl", "curl", "ggplot2", "ggrepel", "maps", "plotly", "stringr", "tm", "wordcloud2", "tidyverse", "RColorBrewer", "ggwordcloud", "viridis", "bookdown", "utils", "leaps", "broom","GGally")
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
        library(x, character.only = TRUE)
    }
})
```

\fancyhead[LR]{}
\pagenumbering{roman}

\newpage
\cleardoublepage
\pagenumbering{arabic}
\fancyhead[L]{Applied Machine Learning and Predictive Modelling: Stroke Data}
\fancyhead[R]{MPM02}


# Introduction
Use case:
We are a smart watch manufacturer working on a new feature for stroke prevention.
We are going to analyze survey data that we plan to ask our users, complementing it with HR (Heart Rate) and CGM (Continuous Glucose Monitoring) data that our product already measures.
We hope that our feature can prevent serious health issues and motivate our users to adopt healthier lifestyles.

# Importing Data
````{r raw_data_view}
stroke_data <- read_csv('./data/healthcare-dataset-stroke-data.csv')

stroke_data

````

```{r}
#testing the effect of non smokers and smokers
count_by_smoke_status <- stroke_data %>% 
  select(smoking_status, stroke) %>%
  group_by(smoking_status, stroke) %>%
  summarise( N = n())

```
```{r}
#boxplot testing the effect of non smokers and smokers
boxplot(N  ~ stroke + smoking_status, data = count_by_smoke_status)
```
```{r}
#LE Testing testing the effect of non smokers and smokers


boxplot(stroke  ~  count_by_smoke_status, data = count_by_smoke_status)
```
```{r}
#Boxplot on avg. glucose_level
```

# Methodology


# Linear Model

# Generalised Linear Model with family set to Poisson

# Generalised Linear Model with family set to Binomial

# Generalised Additive Model

# Neural Network

# Support Vector Machine

# solve an optimisation problem


\newpage
