---
output:
  bookdown::pdf_document2:
    includes:
      in_header: latex/preamble.tex
      before_body: latex/titlepage.tex
    pandoc_args:
    - --csl
    - references/apa.csl
  bookdown::html_document2:
    pandoc_args:
    - --csl
    - references/apa.csl
  bookdown::word_document2:
    pandoc_args:
    - --csl
    - references/apa.csl
toc-title: Table of Contents
bibliography: references/references.bib
link-citations: yes
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \renewcommand{\headrulewidth}{0pt}
- \fancyfoot[C]{}
- \fancyfoot[R]{\thepage}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE)
packages <- c("dplyr","readxl", "curl", "ggplot2", "ggrepel", "maps", "plotly", "stringr", "tm", "wordcloud2", "tidyverse", "RColorBrewer", "ggwordcloud", "viridis", "bookdown", "utils", "leaps", "broom","GGally")
package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
        install.packages(x, dependencies = TRUE)
        library(x, character.only = TRUE)
    }
})
```

\fancyhead[LR]{}
\pagenumbering{roman}

\newpage
\cleardoublepage
\pagenumbering{arabic}
\fancyhead[L]{Applied Machine Learning and Predictive Modelling: Stroke Data}
\fancyhead[R]{MPM02}


# Introduction
Use case:
We are a smart watch manufacturer working on a new feature for stroke prevention.
We are going to analyze survey data that we plan to ask our users, complementing it with HR (Heart Rate) and CGM (Continuous Glucose Monitoring) data that our product already measures.
We hope that our feature can prevent serious health issues and motivate our users to adopt healthier lifestyles.

# Importing Data
````{r raw_data_view}
stroke_data <- read_csv('./data/healthcare-dataset-stroke-data.csv')
stroke_data$bmi <- as.numeric(stroke_data$bmi)
stroke_data$bmi[is.na(stroke_data$bmi)] <- mean(stroke_data$bmi, na.rm = TRUE)
stroke_data$stroke <- as.factor(stroke_data$stroke)

stroke_data

````
## Summaries
```{r}
#testing the effect of non smokers and smokers
count_by_smoke_status <- stroke_data %>% 
  select(smoking_status, stroke) %>%
  group_by(smoking_status, stroke) %>%
  summarise( N = n())

```
```{r}
#testing the effect of work type
 count_by_work_type <- stroke_data %>% 
   select(work_type, stroke) %>% 
   group_by(work_type, stroke) %>%
   summarise( N = n())
```
```{r}
 # testing the effects of residence type
count_by_Residence_type <- stroke_data %>% 
   select(Residence_type, stroke) %>% 
   group_by(Residence_type, stroke) %>%
   summarise( N = n())
```
```{r}
 # testing the effects of gender
count_by_gender <- stroke_data %>% 
   select(gender, stroke) %>% 
   group_by(gender, stroke) %>%
   summarise( N = n())

```
```{r}
 # testing the effects of hypertension
count_by_hypertension <- stroke_data %>% 
   select(hypertension, stroke) %>% 
   group_by(hypertension, stroke) %>%
   summarise( N = n())

```
```{r}
# testing the effects of heart disease
count_by_heart_disease <- stroke_data %>% 
   select(heart_disease, stroke) %>% 
   group_by(heart_disease, stroke) %>%
   summarise( N = n())
```
```{r}
# testing the effects of marriage status
count_by_marriage <- stroke_data %>% 
   select(ever_married, stroke) %>% 
   group_by(ever_married, stroke) %>%
   summarise( N = n())
```







## Scatterplots
```{r scatterplots}
stroke_data %>% 
  ggplot(mapping = aes(x = bmi, y = age, color = stroke)) +
  geom_point() +
  geom_smooth(method = 'lm')

stroke_data %>% 
  filter(stroke == 1) %>%
  ggplot(mapping = aes(x = bmi, y = age, color = smoking_status)) +
  geom_point(method = 'lm') +
  geom_smooth(method = 'lm')

```

## Boxplots

```{r boxplot_gluc}
#Boxplot on avg. glucose_level
ggplot(stroke_data, aes(y = avg_glucose_level)) +
  geom_boxplot()
```

```{r}
#Boxplot on Age 
ggplot(stroke_data, aes(y = age)) +
  geom_boxplot() 
```

```{r}
stroke_by_age <- stroke_data %>%
# dplyr::filter(stroke == 1) %>%
 ggplot(aes(x = stroke,
            y = age)) +
  geom_boxplot()

stroke_by_age
```


```{r}
#Boxplot on BMI
#gplot(stroke_data, aes(y = bmi)) +
#  geom_boxplot()
```

# Methodology


# Linear Model
```{r}
# qplot -> age/glucose level per stroke
library(ggplot2)
 qplot(y = age, x = avg_glucose_level,
        data = stroke_data,
        facets = ~ stroke)
```
```{r}
# fitting models for simple regression model
 lm.stroke <- lm(age ~ avg_glucose_level, data = stroke_data)
 summary(lm.stroke)
```

```


# Generalised Linear Model with family set to Poisson

# Generalised Linear Model with family set to Binomial

# Generalised Additive Model

# Neural Network

# Support Vector Machine

# solve an optimisation problem


\newpage
